Interfacing logic programming and Coq

### 2020-10-28 papers/types2020/cultures.tex

Gentzen's discovery that the key to treating classical and
intuitionistic logics in the same proof system was the invention of
the weakening and contraction structural rules on the right-side of
sequents~\cite{gentzen35}.  This invention lead him away from natural
deduction to multiple-conclusion sequent calculus, a very important
kind of proof that has not been given a satisfactory representation in
type theory.  This same innovation of Gentzen also opened the way to
another key proof-theoretic discovery, that of linear logic
\cite{girard87tcs}.  Similarly, sequent calculus provides an elegant
presentation of linear logic while most treatments of linear logic in
type theory usually only capture linearity in a partial fashion.
\begin{metanote}
  `` partial fashion.'' Isn't that a bit vague? Some references? -am
\end{metanote}

>>>>

Gentzen's discovery that the key to treating classical and
intuitionistic logics in the same proof system was the invention of
the weakening and contraction structural rules on the right-side of
sequents~\cite{gentzen35}.  This invention lead him away from natural
deduction to multiple-conclusion sequent calculus, a very important
kind of proof that has not been given a satisfactory representation in
type theory.  This same innovation of Gentzen also opened the way to
another key proof-theoretic discovery, that of linear logic
\cite{girard87tcs}.  Similarly, sequent calculus provides an elegant
presentation of linear logic while most treatments of linear logic in
type theory have difficulty treating linear negation and the
multiplicative disjunction $\lpar$ and its unit $\bot$.

% DM I don't have a reference but I added more details.
%\begin{metanote}
%  `` partial fashion.'' Isn't that a bit vague? Some references? -am
%\end{metanote}

### 2020-10-28 papers/types2020/pbt.tex

>>>>

\subsection{Implementing the tactic}
The working environment we expect when the tactic is invoked is a proof
environment, where the goal one is trying to prove is the property that the
specification should meet. This means that, after \lsti|intro| has been used to
introduce the relevant hypotheses, we will have a context consisting of
typing information for a bunch of variables that we should divide in two categories:
dependent variables, that appear also in the types of other variables, and non-dependent
variables, that don't appear elsewhere. Obviously, the two are the result of and \lsti|intro| on either
a dependent or a non-dependent product; in our setting, however, this division acquires an additional reading:
since specifications are encoded as pure Horn clauses, were no dependency is involved, non-dependent variables
simply tag the specification clauses whereas dependent variables represent the typing information for the data variables that 
the speficications use.

In order to generate a counterexample, we will need the user to specify which variables of the environment should
be used for generating data and which for executing the specification. In addition to this, the user should 
specify all the certificate information that will guide the data generation phase. The goal of the current environment will
directly provide the property for which we want to find a counterexample. Say we want to generate a counterexample for
some specification represented by \lsti|H1| and \lsti|H2| in the environment, and we want to generate the data for
some variables \lsti|v1| and \lsti|v2| with a height bounded to 10: the final call to the tactic will then 
look like:
\begin{lstlisting}
  elpi pbt 10 (H1 /\ H2) (v1) (v2)
\end{lstlisting}

In order to generate the query to the \lsti|check| predicate, some pre-processing is needed. In particular, we need to
turn the data variables into \lP logic variables when they appear inside a specification, and to generate querys to
\lsti|check| for each of these logic variables in association with the type of the data variable it corresponds to.
In order to code this preprocessing step, we leverage extensively \lP's higher order programming features. The substitutions
are handled with the technique of \lsti|copy| clauses.
